The language of generalization: Probability, vagueness, and interaction

### 2018-04-10 paper/comgen-paper.Rmd
add a note

<!-- Certain linguistic expressions encode generalizations. -->
<!-- Language provides multiple, distinct ways of encoding generalizations. -->
The *language of generalization* covers a diverse swath of natural language expressions.
Generic language conveys generalizations about categories [e.g., "Dogs have four legs"; @Carlson1977; @Cohen1999; @Leslie2007; @Nickel2008] and is the most well-studied case in the language of generalization.^[
  Some writers refer to the all of the *language of generalization* as "generic language".
  In the literature, however, analysis and experiments are often limited to generalizations about categories. 
  We use the more narrow-scoped terms (e.g., generics, habituals, and causals) to highlight the diversity of semantic types being predicated.
]
In contrast to statements about concrete individuals (e.g., "Rufus has four legs"), generic statements refer to inherently unobservable categories (e.g., the category of \textsc{dog}) and convey information that extends beyond the present context, a fact which children as young as 2 appreciate [@Cimpian2008].
Simple events (e.g., "John ran yesterday") can be generalized into habitual sentences (e.g., "John runs"), and even events of complex inferential types such as actual causal events (e.g., "My grandpa drank moonshine and it made him blind") can be described in generalization (e.g., "Drinking moonshine makes you go blind").

@@

It is remarkable that so much is learned from ideas expressed vaguely in words.
Generalizations in language (e.g., "John runs.", "Dogs are friendly.", "Staring at the sun makes you go blind.") are a premier example of how simple statements---statements understood by even the youngest language users---can display complex sensitivities to context.
We have argued that the core meaning of such linguistic expressions can be understood by a vague threshold function that operates over a speaker's inductive beliefs, formalizing by probability.
By precisely articulating the conditions under which generalizations are true or false, we are able to describe how an agent should update its beliefs based on that sentence. 
Thus, in addition to providing a quantitative understanding of the phenomena of the language of generalization, formalizing this language opens up the possibility of building models that learn abstract, generalizable knowledge from language.

As a formal theory of the language of generalization, our framework unifies large swaths of language about distinct semantic types: "Robins lay eggs" (categories), "Sally smokes" (events), "Herb X makes animals sleepy" (causes).
Generalization underlies each of these, and we formalize their meaning by appealing to prevalence, which we define as a predictive probability (as opposed to frequency).
We test this theory by exploring its implications for truth judgments, operationalized as a speaker's decision about whether or not to produce the generalization to naive listener. 
Endorsement, we show, depends in systematic ways on the listener's prior beliefs about the prevalence.
<!-- What these expressions have in common is not surface features but generalization (also referred to as *genericity*). -->
This framework accounts for a number of philosophically puzzling generic statements (Expt. 1), naturally extends to other domains of generalizations (habituals and causals; Expts. 2 \& 3), and provides an avenue for top-down moderators on language use by way of predictive probability (Expt. 2c).
We have provided strict tests of our formal model by both measuring and manipulating listeners' background knowledge as well as comparing our model to a number of alternatives.
<!-- In addition to unifying seemingly disparate parts of language, this paper provides a theoretical framework for asking further questions about genericity, which we outline below. -->

>>>>

<!-- Certain linguistic expressions encode generalizations. -->
<!-- Language provides multiple, distinct ways of encoding generalizations. -->
The *language of generalization* covers a diverse swath of natural language expressions.
Generic language conveys generalizations about categories [e.g., "Dogs have four legs"; @Carlson1977; @Cohen1999; @Leslie2007; @Nickel2008] and is the most well-studied case in the language of generalization.^[
  Some writers refer to the all of the *language of generalization* as "generic language".
  In the literature, however, analysis and experiments are often limited to generalizations about categories. 
  We use the more narrow-scoped terms (e.g., generics, habituals, and causals) to highlight the diversity of semantic types being predicated.
]
In contrast to statements about concrete individuals (e.g., "Rufus has four legs"), generic statements refer to inherently unobservable categories (e.g., the category of \textsc{dog}) and convey information that extends beyond the present context, a fact which children as young as 2 appreciate [@Cimpian2008].
Simple events (e.g., "John ran yesterday") can be generalized into habitual sentences (e.g., "John runs"), and even events of complex inferential types such as actual causal events (e.g., "This fire caused that smoke") can be described in generalization (e.g., "Fire causes smoke").

@@

\mht{let me know where, if anywhere, we should mention "double generalizations" (e.g., generics about habitual properties)}

It is remarkable that so much is learned from ideas expressed vaguely in words.
Generalizations in language (e.g., "John runs.", "Dogs are friendly.", "Fire causes smoke.") are a premier example of how simple statements---statements understood by even the youngest language users---can display complex sensitivities to context.
We have argued that the core meaning of such linguistic expressions can be understood by a vague threshold function that operates over a speaker's inductive beliefs, formalizing by probability.
By precisely articulating the conditions under which generalizations are true or false, we are able to describe how an agent should update its beliefs based on that sentence. 
Thus, in addition to providing a quantitative understanding of the phenomena of the language of generalization, formalizing this language opens up the possibility of building models that learn abstract, generalizable knowledge from language.

As a formal theory of the language of generalization, our framework unifies large swaths of language about distinct semantic types: "Robins lay eggs" (categories), "Sally smokes" (events), "This herb makes zorbs sleepy" (causes).
Generalization underlies each of these, and we formalize their meaning by appealing to prevalence, which we define as a predictive probability (as opposed to frequency).
We test this theory by exploring its implications for truth judgments, operationalized as a speaker's decision about whether or not to produce the generalization to naive listener. 
Endorsement, we show, depends in systematic ways on the listener's prior beliefs about the prevalence.
<!-- What these expressions have in common is not surface features but generalization (also referred to as *genericity*). -->
This framework accounts for a number of philosophically puzzling generic statements (Expt. 1), naturally extends to other domains of generalizations (habituals and causals; Expts. 2 \& 3), and provides an avenue for top-down moderators on language use by way of predictive probability (Expt. 2c).
We have provided strict tests of our formal model by both measuring and manipulating listeners' background knowledge as well as comparing our model to a number of alternatives.
<!-- In addition to unifying seemingly disparate parts of language, this paper provides a theoretical framework for asking further questions about genericity, which we outline below. -->
